### 2012-05-07 3.tex
more writing

\chapter{Static and derived constraints}
\label{turkish}

glossematics

\citet{Chomsky1965} propose

In retrospect, it seems somewhat odd that \citeauthor{Chomsky1965} would coh
It might 
In defining the

The contrast between \emph{blick} and \emph{bnick} is what has been called a ``projection problem'' \citep{Baker1979}

This is not a universal, but rather something English speakers must learn: Moroccan Arabic, for instance, does permit \emph{bn} onsets (e.g., \emph{bni} 'building', \emph{bnat} 'daughters', \emph{bniqa} 'closet').

\citet{Lees1966b,Lees1966a}

This is what we might call a projec
In presenting the components of a 
a term which

by subphonemic details

\citet{Davidson2005,Davidson2006a} shows that excresent schwa is acoustically and articulatorily distinct from "lexical schwa":
\citet{Wilson2011} show taht speakers use this

%English speakers' z^@k is more similar to \emph{scum} than \emph{succumb}

/sC-/ onsets than /s@C-/.

% Davidson 2005: z^@C is more like sC than s@C (ultrasound of articulations)
% Davidson 2006a: lexical and excresant schwa are not the same (acoustics)

% Davidson 2010: production
% Wilson & Davidson 2011:

\citet{Berent2007a}
\citet{Berent2007b}
\citet{Berent2008a}

% Berent et al. 2007
% 598: Highly marked inputs are repaired in perception to abide by the grammatical restrictions of the language
% Exps 1-2: syllable count task (misperception due to linguistic experience; "disyllables were more likely to be perceived as monosyllables if their monosyllabic counterparts had a less marked cluster" in both Russian and English, not what OT predicts)
% Exps 3-4: AB task
% Exps 5-6: priming effect (non-epenthetic and epenethetic things represented the same)
% Berent and Lennertz 2007: (640: If the high rate of monosyllabic errors to unmarked onsets is only due to phonetic failures to encode the input, then it is puzzling why the same onsets also yield the highest rate of accurate response.)
% Berent et al. 2008a:
\citet{Peperkamp2007}
% Peperkamp 2007: during one stage of processing phonology is undone with help from the lexicon...it only takes licit inputs and is irrelevant (p. 633)
\citet{Dupoux1999}
% Dupoux et al. 1999: Japanese speakers still can't distinguish between ebzo and ebuzo in an ABX task using the SAME a or b.
%Elliot Moreton (p.c.)

does not so much as explain the thing itself but rather suggests the

I believe a number of the predictions of this model are wrong
the original and still-compelling 
findings that drove

 motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps. \citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible. \emph{SPE} (p.~3810f.) places this contrast in a system of a system of sequence structure rules that also are used to eliminate lexical redundancy. In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities \citep[][283f.]{Anderson1974}.

This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations. Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion). This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.

the other being the subject

a principle which is one of the two motivations  motivated the introduction of morpheme structure constraints.

In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon does not always require static constraints on possible underlying representations, since such distributions arise by chance and static constraints defined over natural classes contribute little to predicting the contents of the lexicon. 
%This result does not directly bear on the question of whether static restrictions have any synchronic reality.

Yet countless static constraints are statistically supported. Under the null hypothesis advanced in Chapter \ref{MSCs}, the lexical statistics of co-occurrence restrictions are irrelevant: the language acquisition device only attends to derived restrictions and ignores static restrictions regardless of their statistical reliability. On the other hand, many linguists assume that statistical significance provides sufficient conditions for positing synchronically active static constraints. For instance, \citet{Brown2010} writes:

\begin{quotation}
\ldots the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
\end{quotation}

\noindent
A nuanced version of this alternative hypothesis would recognize that there may be other necessary conditions. Any explicit theory of static constraints will have to impose formal restrictions on possible co-occurrence restrictions. The set of possible constraints must respect results in formal learning theory, and may also be, for example, constrained by naturalness biases or formal markedness considerations.

A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements. This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).

Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed. In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.

In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word.

The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements.

The data below, drawn from Turkish, shows that

This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish. Whereas the derived constraint has a robust effect on native speakers' wordlikeness judgements, the static constraint, which is equally statistically valid, has no effect on these judgements. Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.

>>>>

\label{turkish}

%In this chapter, I present evidence that speakers 
%glossematics
%\citet{Chomsky1965} propose

%In retrospect, it seems somewhat odd that \citeauthor{Chomsky1965} would coh
%The contrast between \emph{blick} and \emph{bnick} is what has been called a ``projection problem'' \citep{Baker1979}
%This is not a universal, but rather something English speakers must learn: Moroccan Arabic, for instance, does permit \emph{bn} onsets (e.g., \emph{bni} `building', \emph{bnat} `daughters', \emph{bniqa} `closet').
%\citet{Lees1966b,Lees1966a}

%\citet{Davidson2005,Davidson2006a} shows that this ``excrescent'' schwa is distinct both in acoustics and articulation from ``lexical schwa'': English speakers' [z${^\textrm{ə}}$k] is more similar to \emph{scum} than \emph{succumb}.
%\citet{Wilson2011} show that speakers use this

%\citet{Berent2007a}
%\citet{Berent2007b}
%\citet{Berent2008a}
% 598: Highly marked inputs are repaired in perception to abide by the grammatical restrictions of the language
% Exps 1-2: syllable count task (misperception due to linguistic experience; "disyllables were more likely to be perceived as monosyllables if their monosyllabic counterparts had a less marked cluster" in both Russian and English, not what OT predicts)
% Exps 3-4: AB task
% Exps 5-6: priming effect (non-epenthetic and epenethetic things represented the same)
% Berent and Lennertz 2007: (640: If the high rate of monosyllabic errors to unmarked onsets is only due to phonetic failures to encode the input, then it is puzzling why the same onsets also yield the highest rate of accurate response.)
% Berent et al. 2008a:
%\citet{Peperkamp2007}
% Peperkamp 2007: during one stage of processing phonology is undone with help from the lexicon...it only takes licit inputs and is irrelevant (p. 633)
%\citet{Dupoux1999}
% Dupoux et al. 1999: Japanese speakers still can't distinguish between ebzo and ebuzo in an ABX task using the SAME a or b.
%Elliot Moreton (p.c.)

% motivations for the development of a theory of static co-occurrence restrictions as developed by \citet{SPR}, \citet{Stanley1967} and others is to account for speakers' abilities to distinguish between accidental and structural gaps. \citet{Chomsky1965} propose that a theory of phonology must account for speakers' abilities to distinguish between accidental and structural gaps in their language, citing a contrast between \emph{blick}, which is judged to be a possible but unattested word of English, and \emph{bnick}, which is judged to be impossible. \emph{SPE} (p.~3810f.) places this contrast in a system of a system of sequence structure rules that also are used to eliminate lexical redundancy. In the phonological theory of the era, there is no process that could generate the \emph{blick} $\sim$ \emph{bnick} contrast or numerous other lexical regularities \citep[][283f.]{Anderson1974}.

%This sequence of arguments has already been critiqued in Chapter \ref{1}, where it was shown that prosodic licensing and Stampean occultation places this contrast in core phonology, and that the principles by which speakers recognize impossible words must be stated on surface representations, not underlying representations. Further, \citet[][528f.]{Halle1975} rejects his own principle that lexical entries should be free of all redundancies (see \citealt[][201]{Reiss2003a} and \citealt{Vaux2003} for further discussion). This rejected principle of a redundancy-free lexicon is one of the two principles that motivates Morpheme Structure Constraints; the other is the subject of this and the following chapter.

%a principle which is one of the two motivations  motivated the introduction of morpheme structure constraints.

%In the preceding chapter, it was shown that apparent underrepresentation of certain sequences in the lexicon does not always require static constraints on possible underlying representations, since such distributions arise by chance and static constraints defined over natural classes contribute little to predicting the contents of the lexicon. 
%This result does not directly bear on the question of whether static restrictions have any synchronic reality.

%Yet countless static constraints are statistically supported. Under the null hypothesis advanced in Chapter \ref{MSCs}, the lexical statistics of co-occurrence restrictions are irrelevant: the language acquisition device only attends to derived restrictions and ignores static restrictions regardless of their statistical reliability. On the other hand, many linguists assume that statistical significance provides sufficient conditions for positing synchronically active static constraints. For instance, \citet{Brown2010} writes:

%\begin{quotation}
%\ldots the patterns outlined above are statististically significant. Given this, it stands that these sound patterns should be explained by some linguistic mechanism. \citep[][48]{Brown2010}
%\end{quotation}

%\noindent A nuanced version of this alternative hypothesis would recognize that there may be other necessary conditions. Any explicit theory of static constraints will have to impose formal restrictions on possible co-occurrence restrictions. The set of possible constraints must respect results in formal learning theory, and may also be constrained by markedness considerations, for example.

%A further widespread assumption is that static co-occurrence restrictions validated statistically will be reflected by wordlikeness judgements. This assumption is so deeply entrenched that a number of recent studies of phonotactic learning forgo wordlikeness judgements altogether and instead model the lexical statistics themselves (e.g., \citealt{Coetzee2008a} and \citealt{Anttila2008} on Muna and Arabic, \citealt[][385]{Hayes2008a} on Shona and Wargamay, \citealt{Brown2010} on Gitksan).

%Since the null hypothesis does not countenance statistical significance as evidence for the synchronic reality of a static lexical pattern, such patterns do not adjudicate between the null hypothesis and the alternative, and another source of evidence is needed. In the remainder of Part I, I turn to data from wordlikeness judgement tasks for these purposes.

%In a wordlikeness task, a speaker is presented with nonce words and asked to report such intuitions about the ``possibility'' of a word. 
%The alternative hypothesis, then, is that a statistically reliable co-occurrence restriction will be reflected in native speakers' wordlikeness judgements. 
%The data below, drawn from Turkish, shows that

%This chapter takes the argument one step farther by showing a contrast between static and derived constraints on URs in Turkish. Whereas the derived constraint has a robust effect on native speakers' wordlikeness judgements, the static constraint, which is equally statistically valid, has no effect on these judgements. Anticipating the conclusion, the simplest explanation for the non-effect of this static constraint on wordlikeness judgements is that speakers do not internalize static co-occurrence restrictions at all.

### 2012-05-07 4.1.3.tex
more writing

The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J. \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}). The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}). 
%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
There is extensive motivation for this principle. When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm]. In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [u\lm] to be anomalous \citep{Moreland2009}. This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [ju\lm] also suggests the glide is nuclear. There is also strong external support for this principle. The [ju\lm] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants. The same fusion of [ju\lm] at the expense of the onset is also found in speech errors, e.g., [kju\lm]\emph{mor homponent} for [hju\lm]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}. Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.

@@

%One potential problem with this account is noted by \citet{Kaye1996}, who obseres while [ju\lm] may follow any single tautosyllabic consonant, it never follows branching onsets unless they consist of [s] and a single consonant. 
%This is the only sign that [ju\lm] shows an affiliation for the onset.

@@

The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset. Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}). Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}. Unlike [ju\lm], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}. And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.

%Whereas [Cju\lm] syllables attracts stress, [kwV] syllables do not \citep[][162f.]{Davis1995}.

>>>>

The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J. \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}). The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}). 
%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
There is extensive motivation for this principle. When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː]. In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}. This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [juː] also suggests the glide is nuclear. There is also strong external support for this principle. The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants. The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor homponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}. Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.

@@

%One potential problem with this account is noted by \citet{Kaye1996}, who obseres while [juː] may follow any single tautosyllabic consonant, it never follows branching onsets unless they consist of [s] and a single consonant. 
%This is the only sign that [juː] shows an affiliation for the onset.

@@

The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset. Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}). Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}. Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}. And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.

%Whereas [Cjuː] syllables attracts stress, [kwV] syllables do not \citep[][162f.]{Davis1995}.

### 2012-05-07 4.2.3.tex
more writing

While this procedure can be repeated indefinitely (using the CELEX data, listed in Appendix \ref{A}), consideration of a single characteristic run of the simulation should suffice. The saturation rate of the simulated sample (18.2\%) is very close to the real sample (18.8\%). The slope of the log-rank/log-frequency relationship for the simulated data ($\alpha = -0.557$) is also very close to the observed slope for the CELEX data ($\alpha = -0.588$). And, as shown in Figure \ref{sim}, the two distributions are nearly indistinguishable.

>>>>

While this procedure can be repeated indefinitely 
%(using the CELEX data, listed in Appendix \ref{A}), 
consideration of a single characteristic run of the simulation should suffice. The saturation rate of the simulated sample (18.2\%) is very close to the real sample (18.8\%). The slope of the log-rank/log-frequency relationship for the simulated data ($\alpha = -0.557$) is also very close to the observed slope for the CELEX data ($\alpha = -0.588$). And, as shown in Figure \ref{sim}, the two distributions are nearly indistinguishable.

### 2012-05-07 ack.tex
more writing

So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy. 
More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, whose influence can be seen throughout. 
Gene Buckley, Mark Liberman, and Rolf Noyer, my commitee members, are responsible for creating an environment at Penn where a student could write this admittedly contenious dissertation. 
And though his name appears nowhere else on this document, it is hard to imagine this work could exist without Bill Labov.

My journey in linguistics begins with the mentoring of Jennifer Cole. 
Steve Anderson graciously agreed to serve on my committee. 
Morris Halle told me to never apologize.

I would like to thank my fellow students at Penn, but especially my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody. 
Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document. 
Maryam Bakht kept me centered.

>>>>

So many people contributed to this dissertation that to name them all would compromise the honor of the most worthy. 
More than anyone else, I owe a debt of gratitude to my advisor, Charles Yang, 
who did some much to make this work, and me as a linguist,
into what it is now.
Gene Buckley, Mark Liberman, and Rolf Noyer, my commitee members, are to be thanked for creating an environment at Penn where a student could write this admittedly contenious dissertation. 
And though his name appears nowhere else on this document, it is hard to imagine this work could exist without Bill Labov.

My journey in linguistics begins with the mentoring of Jennifer Cole.
Steve Anderson graciously agreed to serve on my committee.
Morris Halle taught me to speak clearly and never apologize.

While so many of my fellow students at Penn have a hand in this document, 
I would like to especially thank my cohort, Toni Cook, Laurel MacKenzie, and Stefanie Brody.

%Daniel Ezra Johnson fearlessly proofread (\emph{proofreaded}?) the entire document.
%Maryam Bakht kept me centered.
